Probabilistic Policy Violation Detection with IPFWD

### 2017-02-22 paper/main.tex
Update paper, final draft

\author{Philip Nelson and Austin Voecks% <-this % stops a space
\thanks{Supported in part by Dell EMC Isilon through grant and hardware}}

@@

% TODO rewrite
\section{Introduction}

  IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
  chance to accept any packet. The probability of acceptance is updated over
  time and dependent on the current system load. In high performance systems
  this results in an increase in network throughput.

  This approach is based on the premise that firewall performance can be
  improved by reducing the number of rules applied to each packet. Supposing a
  white-list policy, a firewall has to apply every rule to each packet before
  it's denied.  With IPFWD, the system has a chance to accept any packet early
  and skip any further computation. Test results shows this reduces the system
  resources required to handle the same amount of traffic.

  This is a shift in mindset from typical firewalls \cite{networksecurity}.
  Instead of enforcing every part of a security policy all the time, IPFWD
  enforces the policy some of the time and extrapolates from the violations it
  encounters.  This provides insight into the violations that weren't detected.
  For this cost, you gain increased firewall performance and network throughput
  in resource bound systems. It's acceptable to allow a percentage of policy
  violations given that network traffic patterns are often repeated and the
  goal is detection and not immediate prevention. Preventative action may be
  taken later when resource requirements are lower.

  As an example, under heavy load IPFWD may immediately accept 40\% of packets.
  Supposing a port scan was initiated during this time and rules exist to block
  it, 60\% of the port scan would still be rejected and logged. Since the
  probability will fluctuate over time, IPFWD provides information in the IPFW
  logs to show the chance undetected violations occurred for each detected
  violation. IPFWD allows administrators to keep extensive rule sets that fully
  implement their security policy. Instead of having to simplify rule sets to
  increase performance, IPFWD balances policy enforcement and performance
  automatically. Under normal or light load, IPFWD will enforce the entire
  security policy 100\% of the time.

\section{IPFW}

  % Firewall basics
  \subsection{Firewall Basics}

    Firewalls restrict the types of incoming and outgoing network traffic
    according to a security policy. Without a security policy, administrators
    cannot know what to allow and disallow. Typical firewalls enforce the
    policy by blocking packets that do not match the rules. The default action
    for white-list firewalls is to block all traffic, additional rules describe
    exceptions. Blacklist firewalls are the opposite in that by default they
    accept all traffic and additional rules describe packet types that are not
    allowed. Since predicting and blacklisting all possible attacks is
    infeasible, white-list firewall rule sets are the accepted standard and the
    focus of this paper.

@@

    IPFW is a stateful firewall written for FreeBSD which supports both IPv4
    and IPv6. It is comprised of several components: a kernel firewall filter
    rule processor and its integrated packet accounting facility, the logging
    facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
    facility, and an ipstealth facility \cite{freebsdhandbook}.

    IPFW was chosen for it's close integration with the FreeBSD operating
    system, in kernel packet filter, and general performance. Performance is
    dependent on system parameters and circumstances, but IPFW outperforms PF
    in stateful packet filtering \cite{ipfwvspf}. Stateful rules are more
    powerful and flexible than stateless rules, were no session information is
    maintained \cite{networksecurity}.  Additionally, the capabilities of IPFW
    extend beyond packet filtering into source based routing, traffic shaping
    and more.

@@

    This paper is not meant as a comparison between IPFW and PF; the
    functionality provided by IPFWD could be ported to PF and serves as a proof
    of concept for firewalls in general.

@@

  % What does it do?
  IPFWD increases firewall performance by reducing the average amount of work
  it takes to process a packet. Depending on the current system load, IPFWD
  increases or decreases the chance the IPFW will accept a given packet early.
  It does this by adding and maintaining an early rule in the IPFW rule set of
  the following form: \verb|prob 0.000 allow ip from any to any|.

  % How does it work?
  % - Probabilistic packet matching
  IPFWD's behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them. These assumptions will not hold in all environments,
  but we can leverage them for performance gains when they do.

  Since IPFW uses a first-match-wins rule system, this early rule will be
  encountered before the majority of the other rules in the rule set.  The
  probability given in this rule determines the chance that we skip processing
  all further rules, supposing that there hasn't been a match already.

  This is especially valuable for a white-list firewall rule set dealing with
  large amounts of rejected traffic; normally each rejected packet has all
  rules applied to it while IPFW looks for a match.  When no matches are
  encountered, the default rule is employed and the packet is rejected.

  In a typical data center backend network the vast majority of traffic is
  valid and going to be accepted by the firewall's rule set. However, these
  rule sets are often complex and take increasing time to match packets
  depending on the number of rules. For a rule set of length $n$, we can expect
  to a packet to encounter $\frac{n}{2}$ failed matches before matching the
  correct rule and being accepted.

@@

  where $k$ is number of rules before the early acceptance rule and
  $\mathsf{P}(\textit{EA})$ is the probability that the early acceptance rule
  is matched.

  Removing the constant $k$, IPFWD reduces the number of failed matches before
  a successful match by a factor of

@@

  % - Log extension
  While this method increases the speed the system processes packets, it
  creates a chance that packets that would have normally been rejected will get
  through the firewall. We'll refer to these types of packets as invalid for
  the purpose of discussion.

  In order to provide visibility into this process, IPFWD extends the IPFW
  system logs to include information about the early acceptance probability.
  From this information, administrators can extrapolate the probability that
  additional invalid packets made it through the firewall for each that is
  detected. This is discussed further in the \textit{Security Implications}
  section.

@@

        The gains provided by IPFWD are only apparent in systems were network
        throughput is matched by or out performs CPU speed. Systems that can
        already easily handle fully processing each packet will not benefit by
        reducing processing time. Given modern CPU clock speeds, systems with
        network cards slower than 1 Gb/sec will likely not benefit either.

      \item \textbf{Resource Conscious}

        In line with high performance systems, systems with strict resource
        restrictions can benefit by the reduced network related CPU usage.

@@

\section{Security Implications}

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization . At
  the network and host level, this is implemented in part by a firewall rule
  set. A policy violation on an internal network is more serious than an
  external network. Internet facing hosts can expect to be scanned, experience
  network anomalies and be attacked by third party agents semi-regularly.

  However when this happens on an internal network, there is a strong
  possibility something else is going wrong. The usage of IPFWD is predicated
  on the assumption that it's sufficient to be notified of these events so
  further action can be taken later than to always immediately stop them.

  % What does it mean for security if you're ignoring some of the packet?
  Allowing some percentage of network packets through the firewall without
  inspection does mean that attacks that would have normally been blocked could
  succeed. In addition to performance improvements, the following factors help
  to account for this risk:

@@

  Tests were run on three platforms and five different network cards, resulting
  in five different systems. The most important parameters to our tests are
  CPU, NIC, and ENV. Descriptions of each system are given in the tables
  below:\\

@@

    % Explain tools
    Multiple network performance tools were explored but the final results were
    compiled using Netperf. Netperf was chosen because it was readily available
    on all test platforms, provides built-in CPU utilization measurement, and a
    wide range of test types \cite{netperf}.

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
    the target metric, which can be difficult to assess with TCP
    \cite{netperf}.

    Netperf is single threaded, which allows easier comparison of results
    between systems with different numbers of cores. Additionally, all tests
    were run while the system was idle to minimize contention over system
    resources.

    % Experimental method
    All tests were conducted using the following steps:
    \begin{enumerate}
        \item Machine 1
        \begin{itemize}
          \item \texttt{ipfw}
          \item \texttt{netserver}
          \item \texttt{CPU monitor}
        \end{itemize}
        \item Machine 2
        \begin{itemize}
          \item \texttt{ipfw}
          \item \texttt{ipfwd}
          \item \texttt{netperf}
          \item \texttt{CPU monitor}
        \end{itemize}
    \end{enumerate}

    All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules
    depending on the environment and network card configuration. Larger
    production rule sets would benefit even further from IPFWD.

@@

    CPU utilization measurement is the most difficult parameter to measure in
    these tests. There are many factors that can account for variability
    between tests and jitter in the results. The Netperf manual describes these
    challenges and the variety of techniques used on different systems.
    \cite{netperf}

    Netperf was run in single threaded mode, which explains why throughput
    wasn't closer to the maximum throughput for the NIC on each system. The
    FreeBSD network stack allows multithreaded packet processing by utilizing
    multiple queues \cite{freebsdhandbook}. Since we're indirectly testing the
    performance of IPFW through Netperf, there are two different CPU
    measurements to account for: kernel time and user time. Since IPFW runs in
    the kernel, it's CPU time is classified as kernel time.

    Netperf was able to measure CPU utilization on the Virtualized and WWU
    systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
    used. These tests are comparable since both measured user CPU utilization,
    not kernel CPU utilization.

    In systems where the throughput of the NIC exceeds what the CPU can supply,
    CPU utilization is high and throughput is lower than the NIC maximum (eg:
    Virtualised system). Conversely, when CPU availablility exceeds NIC
    throughput, throughput is high and CPU utilization is low (eg: WWU 1G
    system).

@@

    These graphs show the interaction between CPU utilization and network
    throughput for each system with respect to IPFWD's early acceptance
    probability. The dashed lines are CPU utilization, solid is network
    throughput in Mb/second. Higher acceptance probability means a greater
    chance that a given packet is immediately accepted without further
    processing.

    The upward trend in throughput on the x-axis shows that as the early
    acceptance probability increases, so does throughput. The CPU usage shows
    that this throughput is gained without additional CPU utilization, meaning
    the system is spending fewer cycles to process each packet on average.

    The lack of any trend in the WWU 1G graph is expected and addressed in the
    \textit{Discussion} section.

    \begin{center}
      \includegraphics[width=0.45\textwidth]{cpu_isilon40}
      \includegraphics[width=0.45\textwidth]{cpu_isilon10}
      \includegraphics[width=0.45\textwidth]{cpu_wwu1}
      \includegraphics[width=0.45\textwidth]{cpu_virtual10}
    \end{center}

@@

    These graphs show network the ratio of throughput to CPU utilization for
    each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
    presented here is the same as the previous graphs, shown differently. The
    upward trend over the x-axis shows that as the probability of early
    acceptance increases, the system is spending less CPU time to process
    packets on average.

    The dashed line shows the line of best fit for the throughput to CPU
    utilization ratio and shows more clearly the upward trend.

    \begin{center}
      \includegraphics[width=0.45\textwidth]{toc_isilon40}
      \includegraphics[width=0.45\textwidth]{toc_isilon10}
      \includegraphics[width=0.45\textwidth]{toc_wwu1}
      \includegraphics[width=0.45\textwidth]{toc_virtual10}
    \end{center}

  % Difference in results between CPU and throughput bound systems
  \subsection{Discussion}

@@

\section{Related Work}

  % Packet sampling
  Other papers \cite{exploitingpacketsampling, analysisnetflow,
  monitoringpacketsampling} on packet sampling have been focused on detecting
  and classifying threats, not performance enhancement. However, their
  approaches to packet sampling and the implications addressed provided the
  idea for IPFWD.

  % IPFW performance
  The BSD community has produced several in depth investigations into network
  and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
  informed the tests run to measure IPFWD's performance benefits.

@@

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a white-list firewall's rule set would be constructed so that the
  most commonly applied matches are reached earlier rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  is a could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest.

  Other work \cite{redundancydetection} has been done in removing redundant
  rules to increase firewall performance. It's possible that a combination of
  redundancy checking and priority reordering could provide further performance
  gains.

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets 
  Machine learning could be applied to packet samples to more accurately
  determine appropriate bounds on the early acceptance probability. Different
  systems and circumstances may require more or less stringent packet checking.
  Environments with very high confidence in network security might raise the
  lower bound of the acceptance rate to improve performance. Likewise, less
  confident environments may lower the upper bound to ensure an acceptable
  percentage of packets are always checked.

  % have detection of bad traffic reduce acceptance rate
  Future versions of IPFWD could benefit from additional modifiers to the early
  acceptance probability. For instance, the system could reduce the early
  acceptance probability whenever packets are dropped by the firewall under the
  assumption that invalid traffic likely precedes more invalid traffic.

@@

\appendices
\section{Additional Information}
  The source code for IPFWD, additional test results, and development cycle
  information may be found at \texttt{https://???}.

% use section* for acknowledgement
%\section*{Acknowledgment}
% TODO

>>>>

\author{Philip Nelson and Austin Voecks% <-this % stops a space
Western Washington University
\thanks{Supported in part by Dell EMC Isilon through grant and hardware}}

@@

\section{Introduction}

  Firewalls shield networks and hosts from malicious traffic by blocking
  network packets that do not match any of the rules defined in a security
  policy. 
  In high performance and network throughput environments, extensive packet
  processing by firewalls can limit the system's ability to utilize the full
  capacity of network links. This problem is typically addressed by restricting
  the number and complexity of firewall rules, disabling the firewall entirely,
  or sacrificing the additional throughput so all the packet processing can be
  completed.

%  The default action for white-list firewalls is to block all traffic,
%  additional rules describe exceptions. Blacklist firewalls are the opposite,
%  by default they accept all traffic and additional rules describe packet types
%  that are not allowed. Since predicting and blacklisting all possible attacks
%  is not possible in practice, white-list firewall rule sets are the accepted
%  standard and the focus of this paper. Without a security policy,
%  administrators cannot know what to allow and disallow when constructing
%  firewall rules.

    % Packet sampling
    Other papers \cite{exploitingpacketsampling, analysisnetflow,
    monitoringpacketsampling} on packet sampling have been focused on detecting
    and classifying threats, not performance enhancement. However, their
    approaches to packet sampling and the implications addressed provided the
    inspiration for the work done here. The BSD community has produced several in depth investigations into network
    and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
    informed the tests run here.

\section{IPFW \& IPFWD}

  IPFWD is a daemon for FreeBSD that updates an early rule in IPFW (the IP
  Firewall) that has a chance to accept any packet. The probability of early
  acceptance is updated over time and dependent on the current system load. In
  high performance systems this can result in an increase in network
  throughput.

  This approach is based on the premise that firewall performance can be
  improved by reducing the number of rules applied to each packet. Supposing a
  white-list policy, a firewall must apply every rule to a packet before
  it's denied.  With IPFWD, the system has a chance to accept any packet early
  and skip any further computation. Test results show this reduces the 
  resources required to handle the same amount of traffic in some systems.

  This is a shift in mindset from typical firewalls \cite{networksecurity}.
  Instead of enforcing every part of the security policy all the time, IPFWD
  enforces the policy some of the time provides additional information so the
  missed violations can be inferred. For this cost, you gain increased firewall
  performance and network throughput in resource bound systems. IPFWD works
  under the assumption that it's acceptable to allow a percentage of policy
  violations given that network traffic patterns are often repeated and the
  goal is detection, not immediate prevention. Administrative action may be
  taken later when resource requirements are lower.

  As an example, under heavy load IPFWD may immediately accept 40\% of packets.
  Some of those packets may have been malicious. Supposing a port scan was
  initiated during this time and rules exist to block it, at least 60\% of the
  port scan would still be rejected and logged. Since the early acceptance
  probability will fluctuate over time, IPFWD provides information in the IPFW
  logs to show the chance undetected violations occurred for each detected
  violation. IPFWD allows administrators to keep extensive rule sets that fully
  implement their security policy. Instead of having to simplify rule sets to
  increase performance, IPFWD balances policy enforcement and performance
  automatically. Under normal or light load, IPFWD will enforce the entire
  security policy 100\% of the time.

@@

    IPFW is a stateful firewall written for FreeBSD which supports both IPv4
    and IPv6. It is comprised of several components: a kernel firewall filter
    rule processor and its integrated packet accounting facility, a logging
    facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
    facility, and an ipstealth facility \cite{freebsdhandbook}.

    IPFW was chosen for its close integration with the FreeBSD operating
    system, kernel packet filter, and performance. Performance is dependent on
    system parameters and circumstances, but IPFW outperforms the other main
    firewall option for FreeBSD, PF, in stateful packet filtering
    \cite{ipfwvspf}. Stateful rules are more powerful and flexible than
    stateless rules, where no session information is maintained
    \cite{networksecurity}.  Additionally, the capabilities of IPFW extend
    beyond packet filtering into source based routing, traffic shaping and
    more. It's for these reasons that IFPW was chosen over PF.

@@

@@

  % What does it do?
  IPFWD increases firewall performance by reducing the average amount of work
  it takes to process a packet. Depending on the current system load, IPFWD
  increases or decreases the chance the IPFW will accept a given packet early.
  It does this by adding and maintaining an early rule in the IPFW rule set of
  the following form: 
  \begin{center}
      \verb|prob 0.000 allow ip from any to any|
  \end{center}
  which allows any IP packet, incoming or outgoing with a chance equal to the
  probability given.

  % How does it work?
  % - Probabilistic packet matching
  IPFWDs behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them. These assumptions will not hold in all environments,
  but we can leverage them for performance gains when they do.

  Since IPFW uses a first-match-wins rule system, this early rule will be
  encountered before the majority of the other rules in the rule set.  The
  probability given in this rule determines the chance that IPFWD skips
  processing all further rules, supposing that there has not been a match
  already.

  This is especially valuable for a white-list firewall rule set dealing with
  large amounts of rejected traffic; normally each rejected packet has all
  rules applied to it in order to find a match. When no matches are
  encountered, the default rule is employed and the packet is rejected.

  In a typical data center internal network, the vast majority of traffic is
  valid and going to be accepted by the firewalls rule set. However, these rule
  sets are often complex and take increasing time to match packets depending on
  the number of rules. For a rule set of length $n$, we can expect a packet to
  encounter on average $\frac{n}{2}$ failed matches before matching the correct
  rule and being accepted.

@@

  where $k$, a small constant compared to $n$, is number of rules before the
  early acceptance rule and $\mathsf{P}(\textit{EA})$ is the probability that
  the early acceptance rule is matched.

  Simplifying in order notation, IPFWD reduces the number of failed matches before
  a successful match by a factor of

@@

  % - Log extension
  While this method increases the speed the system processes packets, it
  creates a chance that packets that would have normally been rejected will get
  through the firewall.

  In order to provide visibility into this process, IPFWD writes its own logs
  to include information about the early acceptance probability. From this
  information, administrators can extrapolate the probability that additional
  invalid packets made it through the firewall for each that is detected. This
  is discussed further in \ref{securityimplications}.

@@

        The gains provided by IPFWD are only apparent in systems where network
        throughput is matched by or out performs CPU speed. Systems that can
        already easily handle fully processing each packet will not benefit by
        reducing processing time. Given modern CPU clock speeds, systems with
        network cards slower than 1 Gb/sec will likely not benefit from IPFWD.

@@

\section{Security Implications} \label{securityimplications}

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization. At
  the network and host level, this is implemented in part by a firewall rule
  set. A policy violation on an internal network is more serious than an
  external network. Internet facing hosts can expect to be scanned, experience
  network anomalies, and be attacked by third party agents more often than
  internal hosts.

  However, when these events do occur on an internal network, there is a strong
  possibility something else is wrong. Unexpected network activity on a fully
  controlled internal network indicates compromise or misconfiguration, both of
  which are serious problems. The usage of IPFWD is predicated on the
  assumption that it's sufficient to be notified of these events so further
  action can be taken later.

  % What does it mean for security if you're ignoring some of the packet?
  Allowing some percentage of network packets through the firewall without
  inspection does mean that attacks that would have normally been blocked could
  succeed. The following factors help to account for this risk:

@@

  We benchmarked the performance of IPFW in relation to CPU usage and network
  throughput. The goal was to determine if IPFWD reduces the average time IPFW
  takes to process packets. Tests were run on three platforms and four
  different network cards, referred to as four different systems. The most
  important parameters to these tests are CPU, NIC, and ENV. Descriptions of
  each system are given in the tables below:\\

@@

    % Explain tools
    Multiple network performance tools were explored but the final results were
    compiled using Netperf. Netperf generates network traffic between client
    and server instances and measures throughput, errors, and other network
    performance metrics. Netperf was chosen because it was readily available
    on all test platforms, provides built-in CPU utilization measurement, has a
    wide range of test types \cite{netperf}, and is used elsewhere in the BSD
    community for testing.

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance
    was the target metric, which can be difficult to assess with TCP
    \cite{netperf}.

    Netperf is single threaded, which allowed easier comparison of results
    between systems with different numbers of cores. Additionally, all tests
    were run while the system was idle to minimize contention over system
    resources.

    % Experimental method
    All tests were conducted using the following setup:
    \begin{center}
    \begin{tabular}{ | L{10em} | L{10em} | } 
      \hline
      Machine 1 & Machine 2 \\
      \hline
      \hline
      \texttt{ipfw}        & \texttt{ipfw} \\
      \texttt{netserver}   & \texttt{ipfwd} \\
      \texttt{CPU monitor} & \texttt{netperf} \\
                           & \texttt{CPU monitor} \\
      \hline
    \end{tabular} \\
    \end{center}

@@

    CPU utilization measurement was the most difficult parameter to measure in
    these tests. There are many factors that can account for variability
    between tests and jitter in the results. The Netperf manual describes these
    challenges and the variety of techniques used on different systems to
    measure CPU usage \cite{netperf}.

    Netperf was run in single threaded mode, which explains why throughput was
    not closer to the maximum throughput for the NIC on each system. The
    FreeBSD network stack allows multithreaded packet processing by utilizing
    multiple queues meaning the kernel can spread the work done by IPFW over
    multiple cores \cite{freebsdhandbook}. Since we were indirectly testing the
    performance of IPFW through Netperf, there were two different CPU
    measurements to account for: kernel time and user time. Since IPFW runs in
    the kernel, its CPU time is classified as kernel time.

    Netperf was able to measure CPU utilization on the Virtualised and WWU
    systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
    used. \texttt{ps} is a BSD utility that displays information about
    processes, including CPU utilization. We used \texttt{ps} to poll Netperf's
    CPU usage and averaged the output to get the results presented here. These
    tests are comparable since both measured user CPU utilization, not kernel
    CPU utilization. Thus, IPFWs CPU usage was measured indirectly through the
    CPU usage of netperf.

@@

    The graphs in figure \ref{fig:isiloncpu40}, \ref{fig:isiloncpu10},
    \ref{fig:wwucpu1}, and \ref{fig:virtualcpu10} show the interaction between
    CPU utilization and network throughput for each system with respect to
    IPFWDs early acceptance probability. The dashed lines are CPU utilization;
    solid is network throughput in Mb/second. A higher acceptance probability
    means a greater chance that a given packet was immediately accepted without
    further processing.

    The upward trend in throughput on the x-axis shows that as the early
    acceptance probability increases, so does throughput. The CPU usage shows
    that this additional throughput was gained without additional CPU
    utilization, meaning the system was spending fewer cycles to process each
    packet on average.

    The lack of any trend in the WWU 1G graph \ref{fig:wwucpu1} was expected
    and will be addressed in \ref{discussion}.

    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_isilon40}
      \caption{Throughput \& CPU for Isilon 40G system}
      \label{fig:isiloncpu40}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_isilon10}
      \caption{Throughput \& CPU for Isilon 10G system}
      \label{fig:isiloncpu10}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_wwu1}
      \caption{Throughput \& CPU for WWU 1G system}
      \label{fig:wwucpu1}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_virtual10}
      \caption{Throughput \& CPU for Virtualised 10G system}
      \label{fig:virtualcpu10}
    \end{figure}

@@

    The graphs in figures \ref{fig:isilontoc40}, \ref{fig:isilontoc10},
    \ref{fig:wwutoc1}, and \ref{fig:virtualtoc10}  show network the ratio of
    throughput to CPU utilization for each system where $ratio(x) =
    throughput(x) / cpu(x)$. The data being presented here is the same as the
    previous graphs, merely shown differently. The upward trend over the x-axis
    shows that as the probability of early acceptance increases, the system is
    spending less CPU time to process packets on average.

    The dashed line shows the line of best fit for the throughput to CPU
    utilization ratio and shows the upward trend more clearly than the raw
    data.

    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_isilon40}
      \caption{Throughput/CPU for Isilon 40G system}
      \label{fig:isilontoc40}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_isilon10}
      \caption{Throughput/CPU for Isilon 10G system}
      \label{fig:isilontoc10}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_wwu1}
      \caption{Throughput/CPU for WWU 1G system}
      \label{fig:wwutoc1}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_virtual10}
      \caption{Throughput/CPU for Virtualised 10G system}
      \label{fig:virtualtoc10}
    \end{figure}

  % Difference in results between CPU and throughput bound systems
  \section{Discussion} \label{discussion}

@@

    In summary, systems where the throughput of the NIC exceeds what the CPU
    can supply CPU utilization is high and throughput is lower than the NIC
    maximum (eg: the Virtualised system \ref{fig:virtualcpu10}). Conversely,
    when CPU availability exceeds NIC throughput, throughput is high and CPU
    utilization is low (eg: the WWU 1G system \ref{fig:wwucpu1}).

    Additionally, all tests were run with minimal IPFW rule sets in a range of
    30 to 60 rules depending on the environment and network card configuration.
    Larger production rule sets would benefit even further from IPFWD.

@@

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a white-list firewalls rule set would be constructed so that the
  most commonly applied matches are reached sooner rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest. A combination
  of redundancy checking and priority reordering could provide further
  performance gains as well.

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets 
  Further analysis could be done to packet samples in order to more accurately
  determine appropriate bounds on the early acceptance probability. Different
  systems and environments may require more or less stringent packet checking.
  Environments with very high confidence in network security might raise the
  lower bound of the acceptance rate to improve performance. Likewise, less
  confident environments may lower the upper bound to ensure an acceptable
  percentage of packets are always checked.

  % mutli threaded netperf tests
  The tests conducted here were restricted to one core; future work could
  explore the interaction of early acceptance on parallel firewall packet
  processing. Running additional, concurrent netperf tests would show how the
  system behaves when the links are fully saturated.

  % have detection of bad traffic reduce acceptance rate
  Future versions of IPFWD could benefit from additional modifiers to the early
  acceptance probability. For instance, the system could reduce the early
  acceptance probability whenever packets are dropped by the firewall, under
  the assumption that invalid traffic likely precedes more invalid traffic.

@@

\appendices
\section{Additional Information}
  The source code for IPFWD, additional test results, and development cycle
  information may be found at \texttt{https://github.com/Gandalf-/ipfwd}.

\section*{Acknowledgment}

The authors would like to thank Dell EMC Isilon for providing hardware, access
to clusters for testing, and funding for this research. The
resources and the advice on FreeBSD network capabilities/network performance
testing were invaluable. A portion of this work was done during a summer
internship at Dell EMC Isilon.
